# Setting Up The Project Environment {#setup}

## Where Are We - We are **here**!

The importance of reproducible research can be read elsewhere, but I cant stress the importance of consistent file management.

Often work is done in a collaborative manner or on multiple workstations, shared drives etc.

One extremely useful tool in reproducibility and consistency is the **here** package. The here package identifies the root folder and can be used for referencing without the need for absolute referencing.

For example the folder structure for this book is contained in my C: drive and synced with Github, but if I was collaborating with others and used the following absolute reference, it wouldn't work anywhere other than my own computer:

    "C:/Users/nick.dawe/Documents/R Projects Via Git/R Tips - Bookdown Proj"

This is where the **here** package is so useful.

```{r message=FALSE, warning=FALSE, class.source="bg-success"}
library(here)
here()
```
You can use **here** to create various paths as follows -> `here::here("data")`. This code block points to the folder `data` which does not exist (yet).

```{r message=FALSE, warning=FALSE}
library(readr)
dir.create("data")
write.csv(mtcars, here::here("data", "mtcars.csv"))
```
## Standard Folder Structure

I like to standardise my folder structure at the highest level. This allows me to reuse code across projects without worrying about file referencing - especially when using the `here()` function.


```{r}
  here::here()
```

I suggest the following folder names as a minimum - I have not invented this structure myself, but it does form the basic structure for every project I undertake.

```{r set_folder_names}
  folder_names <- c("analysis", # To store RMarkdown documents - I also have a sub folder here called 'figures'
                    "data", # Where the raw data is stored - NEVER store processed data here! 
                    "doc", # Any non R documentation you prepare (.docx reports as needed). I need this, you might not. 
                    "ext", # External information (reference papers, reports etc.)
                    "output", # Processed data
                    "prod", # Production - this is where I store my Shiny apps (usually under a sub folder called "Shiny")
                    "src" # Source code - .R files
                    )
```

Using the vector of `folder_names` we can loop through each item to create the corresponding folder in the home directory. 

```{r create_folders, message=FALSE, warning=FALSE}
for(i in seq_along(folder_names)){
  dir.create(folder_names[i])
}
```

I have written a [package](https://github.com/towananalytics/tatools) containing useful functions for improving project workflow - one of which creates the standard project folder structure above.
